---
title: "Publication_Prep_Script"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```


# Publication Prep Notes: 
-  The relationships being tested here are: SGD --> Settlement, Adult Pocillopora Cover --> Settlement, SGD --> Adult Pocillopora Cover
- Script for figures 2, 4, and supplemental 


### Load libraries 
```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(factoextra)
library(devtools)
library(ggbiplot)
library(MASS)
library(car)
library(performance)
```


### Load Data
```{r}
SettlementDF_Full <- read_csv(here::here("Data", "EnvData_wPocCover.csv"))
## full dataset with pocillopora cover data, env data (nuts, etc), and recruitment counts 
## CommComp --> EnvDataPocCover 

RecruitmentCounts2.1 <- read_csv((here::here("Data", "RecruitmentCounts2.1.csv")))

CoralGenera_Tiles <- read_csv((here::here("Data", "Recruitment_Tiles_Counts.csv")))

SettlementCount_wsalinitypH <- read_csv(here::here("Data", "SettlementCounts_edited_wSALINITYpH.csv"))

WaterChem_withpH <- read_csv(here::here("Data", "pHSlope.csv"))

```


### February, 2026: 
-  Determine acidity diff. in pH units between both sites and ambient SW 
```{r}
exp_pH_diff <- SettlementCount_wsalinitypH %>% 
  dplyr::select(c("treatment_level", "site", "pH")) %>% 
  group_by(site,treatment_level) %>% 
  dplyr::summarize(avg_pH = mean(pH, na.rm=TRUE))

View(exp_pH_diff)

## cabral - varari for:

#### ambient: 
7.977-7.936 ## 0.041

#### low: 
7.966-7.949 ## 0.017

#### med: 
7.948-7.935 ## 0.013

#### high: 
7.927 - 7.920 ## 0.007
```

### Notes about script:
-  Checked if silicate and percent cover are correlated at Varari --> Result = silicate and percent cover are NONLINEARLY correlated at Varari only 
-  For both silicate vs percent cover and silicate vs settlement used "silicate_avg", which are raw values
-  **NOTE:** silicate_avg values were originally calculated from "AllChemData <- read_csv("https://raw.githubusercontent.com/NJS/MooreaSGD_site-selection/main/Data/August2021/Allbiogeochemdata_QC.csv")" in the "RecruitmentTiles_PrelimAnalysis.Rmd" script. Raw silicate values were averaged across tides/days from August 2021. 
-  **NOTE:** Percent cover data was collected from DMB' community surveys --> calculations are shown in "Poc_Cover_Data" script, uses P acuta counts / total counts of biotic organisms per each CowTagID location. Here, values are logged for normalcy of the data

-  Confirming correlation between percent cover and silicate: 
```{r}
##### Confirming correlation between percent cover and silicate #####

## visual test: 

cover_silicate_corrplot <- SettlementDF_Full %>% 
  dplyr::filter(Location=="Varari") %>% 
  ggplot(aes(x = silicate_avg, 
             y = log(percent_cover+1))) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE)

cover_silicate_corrplot


full_varari <- SettlementDF_Full %>% 
  dplyr::filter(Location=="Varari") %>% 
  dplyr::mutate(log_cover = log(percent_cover + 1))

## stats test: 
cor(full_varari$silicate_avg, full_varari$log_cover, use = "complete.obs", method = "pearson")

cor.test(full_varari$silicate_avg, full_varari$percent_cover, method = "pearson")

## nonlinear test: 
nonlinear_test <- lm(log_cover ~ poly(silicate_avg, 2), data = full_varari)
summary(nonlinear_test)


```

-  Why is percent cover only 4%? 
```{r}
#### this because of logging values ####

SettlementDF_Full_test <- SettlementDF_Full %>% 
  dplyr::mutate(log_percentcover = log(percent_cover+1))

#### confirm values of total percent cover + mean/SE at each site ####

SettlementDF_Full_V<- SettlementDF_Full %>% 
  dplyr::filter(Location == "Varari")

SettlementDF_Full_C<- SettlementDF_Full %>% 
  dplyr::filter(Location == "Cabral")


### Varari 
min(SettlementDF_Full_V$percent_cover)
max(SettlementDF_Full_V$percent_cover)
mean(SettlementDF_Full_V$percent_cover)
sd(SettlementDF_Full_V$percent_cover) / sqrt(length(SettlementDF_Full_V$percent_cover))



### Cabral 
min(SettlementDF_Full_C$percent_cover)
max(SettlementDF_Full_C$percent_cover)
mean(SettlementDF_Full_C$percent_cover)
sd(SettlementDF_Full_C$percent_cover) / sqrt(length(SettlementDF_Full_C$percent_cover))

```


-  Check total number of settlers on plates

```{r}

### Varari 
min(SettlementDF_Full_V$sum_total)
max(SettlementDF_Full_V$sum_total)
mean(SettlementDF_Full_V$sum_total)
sd(SettlementDF_Full_V$sum_total) / sqrt(length(SettlementDF_Full_V$sum_total))



### Cabral 
min(SettlementDF_Full_C$sum_total)
max(SettlementDF_Full_C$sum_total)
mean(SettlementDF_Full_C$sum_total)
sd(SettlementDF_Full_C$sum_total) / sqrt(length(SettlementDF_Full_C$sum_total))


```


-  Making new figure (figure 4) for publication that includes 6-subplots: Varari on top (relationship), Cabral on bottom (no relationships) with 1) silicate vs adults, 2) silicate vs settlers, 3) adults vs settlers 

```{r}
##### Varari plots #####

##### settlers vs adults 
adultcover_settlers_V <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% dplyr::filter(Location=="Varari"), 
             aes(x = log(percent_cover+1), 
                 y = sum_total), 
             size = 4, 
             color = "firebrick4") +
  geom_smooth(data = SettlementDF_Full %>% dplyr::filter(Location=="Varari"), 
             aes(x = log(percent_cover+1), 
                 y = sum_total), 
             method = "lm", color="black", alpha=0.5) + 
  theme_classic() + 
  labs(x = "Log(% Cover of Adults)", 
       y= "Total Settlers") + 
  theme(axis.text.x=element_text(size=30), 
        axis.text.y=element_text(size=30), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=40), 
        strip.text=element_text(size=30))

adultcover_settlers_V
ggsave(here::here("Outputs", "Publication", "settlers_adults_V.jpg"), 
       width=11, height=9)

##### silicate vs adults

adultcover_silicate_V <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% dplyr::filter(Location=="Varari"), 
             aes(x = silicate_avg, 
                 y = log(percent_cover+1)), 
             size = 4, 
             color = "firebrick4") +
  geom_smooth(data = SettlementDF_Full %>% dplyr::filter(Location=="Varari"), 
             aes(x = silicate_avg, 
                 y = log(percent_cover+1)), 
             method = "lm", formula = "y~poly(x,2)", color="black", alpha=0.5) + 
  theme_classic() + 
  labs(x = expression(Average~Silicate~(mu*mol~L^-1)), 
       y="Log(% Cover of Adults)") + 
  theme(axis.text.x=element_text(size=30), 
        axis.text.y=element_text(size=30), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=40), 
        strip.text=element_text(size=30))

adultcover_silicate_V
ggsave(here::here("Outputs", "Publication", "silicate_adults_V.jpg"), 
       width=11, height=9)

##### silicate vs settlers 
silicate_settlers_V <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% filter(Location=="Varari"),
             aes(x=silicate_avg, 
                 y=sum_total), 
             size=4, 
             color="firebrick4") +
  geom_smooth(data = SettlementDF_Full %>% filter(Location=="Varari"), 
             aes(x=silicate_avg, 
                  y=sum_total), 
             method = "glm.nb", formula = "y~poly(x,2)", color="black", alpha=0.5) + 
  theme_classic() + 
  labs(x = expression(Average~Silicate~(mu*mol~L^-1)), 
       y="Total Settlers") + 
 theme(axis.text.x=element_text(size=30), 
        axis.text.y=element_text(size=30), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=40), 
        strip.text=element_text(size=30))

silicate_settlers_V
ggsave(here::here("Outputs", "Publication", "silicate_settlers_V.jpg"), 
       width=11, height=9)


##### Cabral plots #####

##### settlers vs adults 

adultcover_settlers_C <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% dplyr::filter(Location=="Cabral"), 
             aes(x = log(percent_cover+1), 
                 y = sum_total), 
             size = 4, 
             color = "goldenrod") +
  geom_smooth(data = SettlementDF_Full %>% dplyr::filter(Location=="Cabral"), 
             aes(x = log(percent_cover+1), 
                 y = sum_total), 
             method = "lm", color="black", alpha=0.5) + 
  theme_classic() + 
  labs(x = "Log(% Cover of Adults)", 
       y="Total Settlers") + 
    theme(axis.text.x=element_text(size=30), 
        axis.text.y=element_text(size=30), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=40), 
        strip.text=element_text(size=30))

adultcover_settlers_C
ggsave(here::here("Outputs", "Publication", "settlers_adults_V.jpg"), 
       width=11, height=9)

#### silicate vs adults 

adultcover_silicate_C <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% dplyr::filter(Location=="Cabral"), 
             aes(x = silicate_avg, 
                 y = log(percent_cover+1)), 
             size = 4, 
             color = "goldenrod") +
 # geom_smooth(data = SettlementDF_Full %>% dplyr::filter(Location=="Cabral"), 
          #   aes(x = silicate_avg, 
             #    y = log(percent_cover+1)), 
            # method = "lm", formula = "y ~ poly(x,2)", color="black", alpha=0.5) + 
  theme_classic() + 
  labs(x = expression(Average~Silicate~(mu*mol~L^-1)), 
       y="Log(% Cover of Adults)") + 
    theme(axis.text.x=element_text(size=30), 
        axis.text.y=element_text(size=30), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=40), 
        strip.text=element_text(size=30))

adultcover_silicate_C
ggsave(here::here("Outputs", "Publication", "silicate_adults_C.jpg"), 
       width=11, height=9)

#### silicate vs settlers
silicate_settlers_C <- ggplot() + 
  geom_point(data = SettlementDF_Full %>% filter(Location=="Cabral"),
             aes(x=silicate_avg, 
                 y=sum_total), 
             size=4, 
             color="goldenrod") +
  theme_classic() + 
  labs(x = expression(Average~Silicate~(mu*mol~L^-1)), 
       y="Total Settlers") + 
    theme(axis.text.x=element_text(size=30), 
        axis.text.y=element_text(size=30), 
        axis.title.x=element_text(size=40), 
        axis.title.y=element_text(size=40), 
        strip.text=element_text(size=30))

silicate_settlers_C
ggsave(here::here("Outputs", "Publication", "silicate_settlers_C.jpg"), 
       width=11, height=9)

```

-  patch them all together in one figure 

```{r}

##### patch plots together #####

full_plot <- (adultcover_silicate_V + adultcover_settlers_V  + silicate_settlers_V) / (adultcover_silicate_C + adultcover_settlers_C + silicate_settlers_C)

full_plot
ggsave(here::here("Outputs", "Publication", "full_plot.jpg"), 
       width=24, height=18)
```


#### run statistical tests for each part of in situ experiments 
-  analyzing site separately to keep with controlled lab experiment + because no pattern with Cabral but yes for Varari

```{r}

###### Varari models #####

# percent cover x settlers 
PocCover_Settlers_V <- glm.nb(sum_total+1 ~ log(percent_cover+1), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Varari"))


anova(PocCover_Settlers_V) ## p = 0.031
summary(PocCover_Settlers_V)
check_model(PocCover_Settlers_V)

# settlers x silicate 
Settlers_SGD_V <- glm.nb(sum_total+1 ~ ((silicate_avg) + I(silicate_avg^2)), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Varari"))


anova(Settlers_SGD_V) ## p = 0.006 for non-linear, lots of collinearity 
summary(Settlers_SGD_V)
check_model(Settlers_SGD_V) # a little bit wonky in dispersion and zero-inflation + homogeniety of variance 

# percent cover x silicate  
## updated on Feb 23rd, 2026 -- changed from glm.nb() to lm() b/c percent cover is not discrete 
PocCover_SGD_V_lm <- lm(log(percent_cover+1) ~ ((silicate_avg) + I(silicate_avg^2)), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Varari"))

anova(PocCover_SGD_V) ## p = 0.029 for non-linear -- this one was WITH glm 
anova(PocCover_SGD_V_lm) ## p = 0.044 for non-linear
check_model(PocCover_SGD_V_lm)

summary(PocCover_SGD_V)
check_model(PocCover_SGD_V) # little bit wonky in dispersion and zero-inflation + homogeniety of variance 


###### Cabral models #####

# percent cover x settlers 
PocCover_Settlers_C <- glm.nb(sum_total+1 ~ log(percent_cover+1), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Cabral"))


anova(PocCover_Settlers_C) ## p = 0.042
summary(PocCover_Settlers_C)
check_model(PocCover_Settlers_C) # little bit wonky in dispersion and zero-inflation

# percent cover x silicate  
## updated on Feb 23rd, 2026 -- changed from glm.nb() to lm() b/c percent cover is not discrete 
PocCover_SGD_C_lm <- lm(log(percent_cover+1) ~ ((silicate_avg) + I(silicate_avg^2)), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Cabral"))


anova(PocCover_SGD_C) ## NOT SIGNIFICANT -- this one was WITH glm 
anova(PocCover_SGD_C_lm) ## STILL NOT SIGNIFICANT 
check_model(PocCover_SGD_V_lm)

summary(PocCover_SGD_C)
check_model(PocCover_SGD_C) # a bit wonky in dispersion and zero-inflation + homogeniety of variance 

# settlers x silicate 
SGD_Settlers_C <- glm.nb(sum_total+1 ~ ((silicate_avg) + I(silicate_avg^2)), data=SettlementDF_Full %>% 
                                  dplyr::filter(Location=="Cabral"))

anova(SGD_Settlers_C) ## NOT SIGNIFICANT
check_model(SGD_Settlers_C)

```

### July 2025: Re-testing Pocillopora Cover with SGD relationship 
```{r}
## plot scatterplot to see general relationship between SGD and coral cover 
SGD_Cover_Relationship <- SettlementDF_Full %>% 
  ggplot(aes(x=silicate_avg, 
             y=log(percent_cover+1), 
             color=Location)) + 
  geom_point(size=3) + 
  theme_bw()

SGD_Cover_Relationship

## model the relationship to test significance 
## only for Varari 

model2 <- lm(log(percent_cover+1) ~ (silicate_avg) + I(silicate_avg^2), data = SettlementDF_Full %>% 
               filter(Location == "Varari"))

anova(model2)

## add site interaction to see how it affects power 
model3 <- lm(log(percent_cover+1) ~ (silicate_avg) + I(silicate_avg^2)*Location, data = SettlementDF_Full)

anova(model3)
```

## re-testing average settlement counts per site

```{r}
## create new dfs to look at means across both sites more easily 
C_tiles <- RecruitmentCounts2.1 %>% 
  filter(Location=="Cabral")

V_tiles <- RecruitmentCounts2.1 %>% 
  filter(Location=="Varari")


# range of settlement counts 
min(C_tiles$sum_total) #0
max(C_tiles$sum_total) #146 

min(V_tiles$sum_total) #2
max(V_tiles$sum_total) #131

# mean and se of settlement counts per site 

## Cabral 
mean(C_tiles$sum_total) #21.65
sd_C <- sd(C_tiles$sum_total, na.rm=TRUE)
se_C <- sd_C/sqrt(20) ##7.87

## Varari 
mean(V_tiles$sum_total) ##33.8
sd_V <- sd(V_tiles$sum_total, na.rm=TRUE)
se_V <- sd_V/sqrt(20) ##8.21

# mean settlement counts per tile 

```


#### Sept 2025: Getting average N and TA values for reef crests 

```{r}

MooreaMarch_full <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/March2022/Allbiogeochemdata_QC_march_fdom2.csv")
MooreaAugust2 <- read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/August2021/Allbiogeochemdata_QC2.csv")


MooreaAugust2_edit <- MooreaAugust2 %>% 
  select(-c(,2:7,10:14,16,24:30))
MooreaMarch_edit <- MooreaMarch_full %>% 
  select(-c(,2:7,10:14,16,24:30))

MooreaAugust_addedOffshore <- MooreaMarch_edit %>%  
  full_join(MooreaAugust2_edit)

nut_dfedit <- MooreaAugust_addedOffshore %>% 
  dplyr::filter(Plate_Seep == "Offshore") %>% 
  dplyr::select(c(1,3,5,6,10)) %>% 
  dplyr::filter(Date != "3/26/2022"&Date != "3/27/2022"&Date != "3/28/2022")

#####
V_nuts <- nut_dfedit %>% 
  dplyr::filter(Location=="Varari") %>% 
  dplyr::summarize(avgTA=mean(TA), 
                   sdTA = sd(TA),
                   seTA = sdTA / sqrt(length(TA)),
                   avgN=mean(NN_umolL, na.rm=TRUE),
                   sdN = sd(NN_umolL, na.rm=TRUE), 
                   seN = sdN/sqrt(length(NN_umolL)))

C_nuts <- nut_dfedit %>% 
  dplyr::filter(Location=="Cabral") %>% 
  dplyr::summarize(avgTA=mean(TA), 
                   sdTA = sd(TA),
                   seTA = sdTA / sqrt(length(TA)),
                   avgN=mean(NN_umolL, na.rm=TRUE),
                   sdN = sd(NN_umolL, na.rm=TRUE), 
                   seN = sdN/sqrt(length(NN_umolL)))
```

## Figure 2: ggridges 

```{r}
## re-arrange data for better format 

SettlementCount_salinitypH_pivot <- SettlementCount_wsalinitypH %>% 
  dplyr::select(!c("swimming":"spawning_day", "notes")) %>% 
  pivot_longer(cols = silicate:NN|salinity:pH, names_to= "parameters", values_to = "values")

# ggridges plot 

ggridges_plot <- SettlementCount_salinitypH_pivot %>% 
  mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = values, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
  facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot

##################
## silicate
##################
ggridges_plot_silicate <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = silicate, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_silicate

##################
## nitrates
##################
ggridges_plot_NN <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = NN, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_NN

##################
## salinity
##################
ggridges_plot_salinity <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = salinity, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_salinity


##################
## pH
##################

ggridges_plot_pH <- SettlementCount_wsalinitypH %>% 
 # mutate(parameters = factor(parameters, levels = c("silicate", "NN", "salinity", "pH"))) %>% 
  ggplot(aes(x = pH, 
             y = treatment_level, 
             fill = site)) +
  geom_density_ridges(alpha=0.7, scale = 3, rel_min_height = 0.01) +
  scale_fill_manual(values=c("darkgoldenrod2", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_ipsum() +
 # facet_wrap(~parameters, scales="free_x") +
  scale_y_discrete(limits=c("High", "Medium", "Low", "Ambient")) +
    theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20), 
        panel.spacing = unit(0.1, "lines"),
        strip.text.x = element_text(size = 20)) + 
  labs(x = "", 
       y = "") 

ggridges_plot_pH


## add together 
ggridges_plot_silicate + ggridges_plot_NN + ggridges_plot_salinity + ggridges_plot_pH


ggsave(here::here("Outputs", "Publication", "GGridges.jpg"),
       width=12, height=10)
```


## Figure 2: endmember plots 

### Create table with environmental data 
```{r}
###############################
## in case someone wants to see all the raw values but kept out of publication otherwise 
###############################
modTable <- WaterChem_withpH %>%  
  dplyr::select(!c("TrisCalDate":"TTris", "note", "sampling_period", "light", "experiment_number", "orion_pH", "mV", "mVTris")) %>%
  drop_na(TA) %>% 
  mutate(treatment_level = if_else(treatment_level == "Fresh", "Raw GW", treatment_level)) %>% 
  rename(Phosphate = phosphate, 
         Silicate = silicate, 
         Ammonia = ammonia,
         "Nitrate+Nitrite" = NN, 
         "Total Alkalinity" = TA, 
         "Treatment Rep" = treatmentID_rep, 
         "SGD Level"= treatment_level, 
         "Time of Measurement" = time, 
         Date = date, 
         Temperature = temp, 
         Salinity = salinity) 

modTable

###############################
## Average for treatments so table is reasonable size 
###############################
modTable2 <- WaterChem_withpH %>%  
  dplyr::select(!c("TrisCalDate", "note", "sampling_period", "light", "experiment_number", "orion_pH")) %>% 
  drop_na(TA) %>% 
  mutate(treatment_level = if_else(treatment_level == "Fresh", "Raw GW", treatment_level)) %>% 
  group_by(Location, treatment_level) %>% ## removed time and just included date so it is even more manageable -- took out date
  summarize("AvgTemp" = mean(temp), 
            "AvgPhosphate" = mean(phosphate), 
            "AvgSilicate" = mean(silicate), 
            "AvgNitrateNitrite" = mean(NN), 
           # "AvgTotalAlkalinity" = mean(TA), 
            "AvgSalinity" = mean(salinity), 
            "AvgpH" = mean(pH)) %>% 
  mutate(AvgSalinity = round(AvgSalinity, 2), 
       #  AvgTotalAlkalinity = round(AvgTotalAlkalinity, 2), 
         AvgNitrateNitrite = round(AvgNitrateNitrite, 2), 
         AvgSilicate = round(AvgSilicate, 2), 
         AvgPhosphate = round(AvgPhosphate, 2), 
         AvgTemp = round(AvgTemp, 2), 
         AvgpH = round(AvgpH, 2)) %>% 
 # mutate(treatment_level = factor(treatment_level, levels = c("Low", "Medium", "High"))) %>%
  rename("SGD Level"= treatment_level, 
         #"Time of Measurement" = time, 
        # Date = date, 
         "Avg Temperature" = AvgTemp, 
         "Avg Phosphate" = AvgPhosphate, 
         "Avg Silicate" = AvgSilicate, 
         "Avg Nitrate+Nitrite" = AvgNitrateNitrite, 
       #  "Avg Total Alkalinity" = AvgTotalAlkalinity, 
         "Avg Salinity" = AvgSalinity, 
         "Avg pH" = AvgpH) 

modTable2

modTable2 <- modTable2 %>%
  mutate('SGD Level' = factor('SGD Level', levels = c('Raw GW', "Ambient", "Low", "Medium", "High")))  

# Create the table with the manually ordered SGD Level
EnvData_Table <- modTable2 %>%
  arrange('SGD Level') %>%  # Arrange by the ordered factor
  kbl() %>%
  kable_classic(html_font = "Times New Roman", font_size = 14) %>%
  row_spec(0, italic = TRUE, bold = TRUE) %>%
  pack_rows("Cabral", 1, 5) %>%
  pack_rows("Varari", 6, 10) %>%
  column_spec(1:6, background = "white")

```

### Re-design table with Kable 
```{r}
 
EnvData_Table <- modTable2 %>%
  arrange('SGD Level') %>%  # Arrange by the ordered factor
  kbl() %>%
  kable_classic(html_font = "Times New Roman", font_size = 14) %>%
  row_spec(0, italic = TRUE, bold = TRUE) %>%
  pack_rows("Cabral", 1, 5) %>%
  pack_rows("Varari", 6, 10) %>%
  column_spec(1:6, background = "white")

EnvData_Table

######################
## save table 
######################
EnvData_Table %>% 
  as_image(file = here("Outputs", "Publication", "EnvData_Table.jpg"))



```

### make plots with averaged endmember groundwater data 
```{r}
modTable2_edit <- modTable2 %>% 
  rename(treatment_level ="SGD Level", 
         AvgTemp = "Avg Temperature",
         AvgPhosphate ="Avg Phosphate" , 
         AvgSilicate = "Avg Silicate", 
         AvgNitrateNitrite = "Avg Nitrate+Nitrite", 
         #AvgTotalAlkalinity = "Avg Total Alkalinity", 
         AvgSalinity = "Avg Salinity", 
         AvgpH = "Avg pH") %>% 
 # dplyr::filter(treatment_level=="Raw GW"|treatment_level=="Ambient") %>% 
  group_by(Location, treatment_level) %>% 
  summarize(meantemp = mean(AvgTemp), 
            meanphos = mean(AvgPhosphate), 
            meansilicate = mean(AvgSilicate), 
            meanNN = mean(AvgNitrateNitrite),
         #   meanTA = mean(AvgTotalAlkalinity),
            meanpH = mean(AvgpH), 
            meansal = mean(AvgSalinity), 
            setemp = sd(AvgTemp) / sqrt(n()), 
            sephos = sd(AvgPhosphate) / sqrt(n()), 
            sesilicate = sd(AvgSilicate) / sqrt(n()),
            seNN = sd(AvgNitrateNitrite) / sqrt(n()), 
         #   seTA = sd(AvgTotalAlkalinity) / sqrt(n()), 
            sesal = sd(AvgSalinity) / sqrt(n()), 
            sepH = sd(AvgpH) / sqrt(n())) # %>% 
 # pivot_longer(cols = "meantemp":"meansal", names_to = "avg_params", values_to = "avg_values")

################
## plot 
################

## pH
endmember_plot_pH <- modTable2_edit %>% 
  filter(treatment_level=="Raw GW") %>% 
  ggplot(aes(x=Location, 
             y=meanpH, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanpH - sepH, 
                    ymax = meanpH + sepH), 
                width = 0.2, position = position_dodge(0.7)) + 
  labs(x = "", 
      y = "") + 
 coord_cartesian(ylim = c(7, 8)) +
 # scale_y_continuous(limits = c(7, 8)) +
       # y = expression(Average~pH) + 
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_pH


## NN
endmember_plot_NN <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meanNN, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanNN - seNN, 
                    ymax = meanNN + seNN), 
                width = 0.2, position = position_dodge(0.7)) + 
     labs(x = "", 
          y = "") + 
       # y = expression(Average~Nitrate~Nitrite(mu*mol~L^-1))) + 
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_NN

## silicate 

endmember_plot_sil <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meansilicate, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meansilicate - sesilicate, 
                    ymax = meansilicate + sesilicate), 
                width = 0.2, position = position_dodge(0.7)) + 
     labs(x = "", 
          y = "") + 
       # y = expression(Average~Silicate~(mu*mol~L^-1))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_sil

## salinity 

endmember_plot_sal <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meansal, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"),
                     guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meansal - sesal, 
                    ymax = meansal + sesal), 
                width = 0.2, position = position_dodge(0.7)) + 
     labs(x = "", 
          y = "") + 
        #y = expression(Average~Salinity~(psu))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_sal


## phosphate 

endmember_plot_phosphate <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meanphos, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanphos - sephos, 
                    ymax = meanphos + sephos), 
                width = 0.2, position = position_dodge(0.7)) + 
   labs(x = "", 
        y = expression(Average~Phosphate~(mu*mol~L^-1))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_phosphate

## TA 

endmember_plot_TA <- modTable2_edit %>% 
  ggplot(aes(x=Location, 
             y=meanTA, 
             fill=Location)) + 
  geom_col(position="dodge") + 
#  facet_wrap(~avg_params, scales="free_y") + 
  scale_fill_manual(values=c("goldenrod", "firebrick4"), 
                    limits=c("Cabral", "Varari"), 
                    guide="none") + 
  theme_classic() + 
  geom_errorbar(aes(ymin = meanTA - seTA, 
                    ymax = meanTA + seTA), 
                width = 0.2, position = position_dodge(0.7)) + 
    labs(x = "", 
        y = expression(Average~TA~(mu*mol~L^-1))) + 
   theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))

endmember_plot_TA

######################################
## add them all together 
######################################

endmember_plots <-  endmember_plot_sil + endmember_plot_NN + endmember_plot_sal + endmember_plot_pH

endmember_plots
ggsave(here::here("Outputs", "Publication", "endmember_plots.jpg"), 
       width=12, height=8)
```


## Supplemental Figure: 

```{r}
## tidy data in way that makes sense for plot 

CoralGenera_Tiles_edit <- CoralGenera_Tiles %>% 
  dplyr::select(!"Notes") %>% 
  pivot_longer(cols="Pocilloporidae":"Other", names_to = "coral_genera" , values_to= "counts") %>% 
  group_by(CowTagID, Location, coral_genera) %>%  
  summarize(genera_total=sum(counts)) %>% 
  ungroup() %>% 
  group_by(CowTagID, Location) %>%  
  mutate(sum_total=sum(genera_total)) %>% 
  mutate(percentage = (genera_total/sum_total)*100) %>% 
  filter(CowTagID!="C19") %>% 
  ungroup() %>% 
  group_by(coral_genera, Location) %>%
  mutate(mean_percentage = mean(percentage),
         se = sd(percentage) / sqrt(n()))  # Standard error calculation

#######################################
## for stacked bar plot - no averages 
########################################

## calculate total at Varari and at Cabral 
      ## 433 at Cabral and 676 at Varari 

CoralGenera_Tiles_editV <- CoralGenera_Tiles %>% 
  dplyr::select(!"Notes") %>%
  filter(Location=="Varari") %>% 
  pivot_longer(cols="Pocilloporidae":"Other", names_to = "coral_genera" , values_to= "counts") %>% 
  group_by(Location, coral_genera) %>% 
  summarize(genera_total=sum(counts), 
            percent = (genera_total/676)*100)

CoralGenera_Tiles_editC <- CoralGenera_Tiles %>% 
  dplyr::select(!"Notes") %>%
  filter(Location=="Cabral") %>% 
  pivot_longer(cols="Pocilloporidae":"Other", names_to = "coral_genera" , values_to= "counts") %>% 
  group_by(Location, coral_genera) %>% 
  summarize(genera_total=sum(counts), 
            percent = (genera_total/433)*100)


CoralGenera_Tiles_edit2 <- CoralGenera_Tiles_editC %>% 
  bind_rows(CoralGenera_Tiles_editV)
  
########################################
## plot percent total of the genera
########################################

coral_generaplot1 <- CoralGenera_Tiles_edit2 %>% 
  mutate(grp = factor(coral_genera, levels = c("Other","Acroporidae","Poritidae","Pocilloporidae"))) %>% # put the levels in an order I want for plotting
  ggplot(aes(x=Location, 
             y=percent, 
             fill=grp)) + 
  geom_col() + 
  #geom_errorbar(aes(ymin = mean_percentage - se, 
                  #  ymax = mean_percentage + se), 
               # width = 0.2, position = position_dodge(0.7)) +  # Add error bars
  theme_classic() + 
  scale_fill_manual(values=c("gray", "coral2", "chocolate4", "burlywood"), 
                    limits=c("Other", "Acroporidae", "Poritidae", "Pocilloporidae")) +
  labs(x = "Location", 
       y= "Percentage of Total Settlers", 
       fill= "Coral Family") + 
  theme(axis.text.x=element_text(size=20), 
        axis.text.y=element_text(size=20), 
        axis.title.x=element_text(size=20), 
        axis.title.y=element_text(size=20), 
        legend.text=element_text(size=20),  # Adjust legend text size
        legend.title=element_text(size=20), # Adjust legend title size
        strip.text=element_text(size=20))
  
coral_generaplot1
ggsave(here::here("Outputs", "Publication", "coral_genera_tiles.jpg"), 
       width=10, height=7)
```







